\documentclass[11pt]{article}
\usepackage{enumitem}
\usepackage{latexsym}
\usepackage{amsfonts}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{xcolor}

\usepackage{tikz}
\usetikzlibrary{arrows}
\usetikzlibrary{automata}

\setlength{\textheight}{8.5in}
\setlength{\textwidth}{6.0in}
\setlength{\headheight}{0in}
\addtolength{\topmargin}{-.5in}
\addtolength{\oddsidemargin}{-.5in}

\input{../lecture_notes/preamble.tex}

\newcommand{\solution}[1]{\paragraph{Solution}  }
\newcommand{\bl}[1]{\textcolor{blue}{#1}}
\newcommand{\rd}[1]{\textcolor{red}{#1}}

\tikzset{
->, % makes the edges directed
>=stealth, % makes the arrow heads bold
node distance=4cm, % specifies the minimum distance between two nodes. Change if necessary.
every state/.style={thick, fill=gray!10}, % sets the properties for each ’state’ node
initial text=$ $, % sets the text that appears on the start arrow
}

\begin{document}
\hw{2: Context-Free Languages}{1/28/2023}{Rishav Bhagat}{Due:2/17/2023}

You should submit a typeset or \emph{neatly} written pdf on Gradescope.  The grading TA should not have to struggle to read what you've written; if your handwriting is hard to decipher, you will be asked to typeset your future assignments. Five bonus points if you use \LaTeX, and our template. You may collaborate with other students, but any written work should be your own.

%CFGs, PDAs, pumping them
\begin{enumerate}

    \item Give a CFG for valid regular expressions over the alphabet $\Sigma = \{a,b\}$. The alphabet of your grammar should be over $\{a,b,(,), \cup, ^*, \lambda, \emptyset\}$. You may insert more parenthesis than necessary. Here we shall use $\lambda$ to represent the base-case regular expression symbolizing the set only containing the empty string. Use $\varepsilon$ not as part of the regular expression, but if necessary, as part of the right-hand-side of productions of the grammar to satisfy $(V \cup \Sigma)^*$. You must generate all syntactically valid regular expressions.

    \solution{1} 

    $$ S \rightarrow (S^*) | (S \cup S) | (SS) | a | b | \emptyset | \lambda $$
    
    \item Give a CFG for the language $\{a^ib^j~|~x,y \in \mathbb{Z}, $ $x+y = i$ and $4x+3y = j\}$. Then convert it to CNF form. Show your work.

    \solution{2} CFG for $\{a^ib^j~|~x,y \in \mathbb{Z}, $ $x+y = i$ and $4x+3y = j\}$:

    $$ S \rightarrow aSbbbb | aSbbb | \epsilon $$

    Now, we must convert it to a CNF, using the following steps from lecture:

    a) Add a new start state

    $$ S \rightarrow T $$

    $$ T \rightarrow aTbbbb | aTbbb | \epsilon $$

    b) Delete and patch all $\epsilon$ rules

    $$ S \rightarrow T | \epsilon $$

    $$ T \rightarrow aTbbbb | aTbbb | abbbb | abbb $$

    c) Replace all unit rules

    $$ S \rightarrow aTbbbb | aTbbb | abbbb | abbb | \epsilon $$

    $$ T \rightarrow aTbbbb | aTbbb | abbbb | abbb $$

    d) Creates rules of the form $A \rightarrow u_1 u_2$

    $$ S \rightarrow aT_4 | aT_3 | aB_4 | aB_3 | \epsilon $$

    $$ T \rightarrow aT_4 | aT_3 | aB_4 | aB_3 $$

    $$ T_4 \rightarrow T B_4 $$
    $$ T_3 \rightarrow T B_3 $$

    $$ B_4 \rightarrow bB_3 $$
    $$ B_3 \rightarrow bB_2 $$
    $$ B_2 \rightarrow bB $$
    $$ B \rightarrow b $$

    e) $\forall a \in \Sigma$ replace $a$ by $A$ and add rule $A \rightarrow a$

    $$ S \rightarrow AT_4 | AT_3 | AB_4 | AB_3 | \epsilon $$

    $$ T \rightarrow AT_4 | AT_3 | AB_4 | AB_3 $$

    $$ T_4 \rightarrow T B_4 $$
    $$ T_3 \rightarrow T B_3 $$

    $$ B_4 \rightarrow BB_3 $$
    $$ B_3 \rightarrow BB_2 $$
    $$ B_2 \rightarrow BB $$

    $$ A \rightarrow a$$
    $$ B \rightarrow b $$

    The above CNF is equivalent to the CFG produced.

    \newpage

    \item Give a PDA for the language $\{w \in \Sigma^* ~|~ w \text{ has even length and is not a palindrome}\}$

    \begin{figure}[ht]
        \centering
        \begin{tikzpicture}
            \node[state, initial] (q0) {$q_0$};
            \node[state, right of=q0] (q1) {$q_1$};
            \node[state, below of=q1] (q2) {$q_2$};
            \node[state, below of=q2] (q3) {$q_3$};
            \node[state, right of=q3, below of=q2, accepting] (q4) {$q_4$};
            
            \draw   (q0) edge[above] node{$\epsilon, \epsilon \rightarrow \$$} (q1)
                    (q1) edge[loop right, right] node[text width=2cm, align=left]{$a, \epsilon \rightarrow a \\ b, \epsilon \rightarrow b$} (q1)

                    (q1) edge[right] node{$\epsilon, \epsilon \rightarrow \epsilon$} (q2)
                    (q2) edge[loop right, right] node[text width=2cm, align=left]{$a, a \rightarrow \epsilon \\ b, b \rightarrow \epsilon$} (q2)

                    (q2) edge[right] node[text width=2cm, align=left]{$a, b \rightarrow \epsilon \\ b, a \rightarrow \epsilon$} (q3)
                    (q3) edge[loop left, left] node[text width=2cm, align=right]{$a, a \rightarrow \epsilon \\ b, b \rightarrow \epsilon \\ b, a \rightarrow \epsilon \\ a, b \rightarrow \epsilon$} (q3)

                    (q3) edge[above] node{$\epsilon, \$ \rightarrow \epsilon$} (q4)

                    ;
        \end{tikzpicture}
    \end{figure}

    \newpage
    
    \item Prove $\{x_1\star x_2\star...\star x_t ~|~ t > 1, x_i \in \{a,b\}^*, \text{ and } \exists x_i,x_j \text{ with } i \neq j \text{ but } x_i = x_j \}$ is not context-free.

    \solution{4} Assume, to the contrary, that $L = \{x_1\star x_2\star...\star x_t ~|~ t > 1, x_i \in \{a,b\}^*, \text{ and } \exists x_i,x_j \text{ with } i \neq j \text{ but } x_i = x_j \} $ is a context free language (CFL) with pumping length $p$.

    Notice that $s = a^pb^p\star a^pb^p$ is a string within the $L$ with length $|s| \geq p$. Now, we consider all subdivisions of $s$, $s=uvxyz$ such that $|vxy|\leq p$ and $|vy| > 0$.

    There are three cases of where the $\star$ can be: \\
        1. $\star \notin vxy$ \\
        2. $\star \in x$ \\
        3. $\star \in vy$. 

    Notationally, let's say that $a_1$ is the sub-string of $s$ before (not including) the $\star$ and $a_2$ is the sub-string of $s$ after (not including) the $\star$. Also, let $r = uv^ixy^iz$ be the pumped string given an $i$. Additionally, let $b_1$ be the sub-string of the pumped string $r$ before (not including) the $\star$ and $b_2$ be the sub-string of the pumped string $r$ after (not including) the $\star$.

    \underline{Case 1}: $\star \notin vxy$:

    Since $vxy$ is only either (but not both) fully contained in $a_1$ or $a_2$, any pumping will break the symmetry and $b_1 \neq b_2$ in the resulting pumped string. This would imply that $r$ is not in $L$ for any $i \neq 1$. Thus, we can choose $i=2$ to contradict the pumping lemma for this case.

    \underline{Case 2}: $\star \in x$:

    Then, $v$ only can contain $b$'s and $y$ can only contain $a$'s since $|vxy| \leq p$ and there are p $b$'s and $a$'s. Thus, again, setting any $i \neq 1$ value will break the symmetry and $b_1 \new b_2$ in the resulting pumped string, meaning that $r \notin L$. Thus, we can choose $i=2$ to contradict the pumping lemma for this case. 

    \underline{Case 3}: $\star \in vy$:

    If the star is somewhere in $v$ or $y$, we can pump down and the star will no longer be present in the resulting string $r$. Thus, when $i=0$, $\star \notin r$, which means that $t=1$ for $r$, breaking the condition that $t > 1$. Thus, $r \notin L$, contradicting the pumping lemma for this last case.

    Since the pumping lemma has been contradicted for all possible subdivisions of $s$ under the conditions mentioned above, $L$ cannot be a context-free language.

    
\end{enumerate}
\end{document}
